---
title: "Assessing Epidemiological Level of Risk (ELR)"
author: "Thibaut Jombart"
date: '`r format(Sys.time(), "%A %d %B %Y")`'
output: 
  html_document:
    toc: TRUE
    toc_depth: 4
    toc_float: TRUE
    toc_collapse: FALSE
    number_sections: TRUE
    highlight: pygments
    theme: spacelab
    code_folding: hide
    <!-- df_print: paged -->
    css: !expr here::here('css', 'style.css')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 70,
                      warning = TRUE,
                      message = TRUE)
```

```{r eval = FALSE, echo = FALSE}

rmarkdown::render('elr_synthesis.Rmd')

```


# Prep work {.tabset .tabset-fade .tabset-pills}

## Scripts

```{r warnings = FALSE}

rfextras::load_scripts()

```


## Packages

All CRAN packages can be installed by calling `reportfactory::install_deps()`.
Package installs on remotes are now handled in a separate script
`remote_packages.R` which is loaded alongside other scripts in `scripts/` using:


```{r}

pacman::p_load(
  tidyverse,
  rio,
  ggrepel
)

# load separately github-only packages
library(trendbreaker)

```



## Load the data

We load the `$df_dynamics` components of the list of results stored for each WHO
region, output by `regional_analyses.Rmd`:

```{r }

all_files <- list.files(here::here("asmodee_outputs"),
                        pattern = "latest.rds",
                        recursive = TRUE,
                        full.names = TRUE)

all_content <- map(all_files, rio::import)
regions <- all_files %>%
  gsub("[[:lower:][:punct:]]", "", .)

all_df <- all_content %>%
  map(list("df_dynamics")) %>%
  setNames(regions)

all_elr_extras <- all_content %>%
  map_df(list("elr_extras"))


```

Add region data and assemble into a single `data.frame`, keeping only the
current day:

```{r }


for (i in seq_along(regions)) {
  all_df[[i]] <- mutate(all_df[[i]], who_region = regions[i])
}

x_all <- all_df %>%
  bind_rows()

x <- x_all %>%
  filter(last_day == 0)

```


## Load VOI data

```{r }

dat_voi <- find_latest("proportions_voi", where = here::here("data")) %>%
  rio::import() %>%
  pivot_wider(names_from = quantile, values_from = value)

dat_voi


```


------------------------------------

# Generate pre-classification {.tabset .tabset-fade .tabset-pills}

## Classification algorithm

<img src = "https://raw.githubusercontent.com/whocov/trend_analysis_public/main/docs/images/decision_CoC.png" width = "70%">

We define the following classes for levels of concerns:

* not listed
* *low*: average incidence *and* no decline
* *mid*: high incidence *or* fast growth, **or** average incidence *and* growth
* *high*: high/very high incidence *and* growing, **or** low incidence *and* growing fast
  very fast
* *very high*: high/very high incidence *and* growing fast/very fast

We use the following classification rules, depending on the values of the growth
rate (*r*), current incidence (as % of previous peak) and signs of acceleration
(ASMODEE net increases over last week):

* if incid < 40%:
    + if r > high_r -> **M**
    + else if r > 0.02 -> **L**
	+ else -> Minimal
* if 40% <= incid <= 80%:
    + if r > high_r -> **H**
    + else if r > 0.02 -> **M**
	+ else if r is not < 0 -> **L**
    + else -> **Minimal**
* if incid > 80%:
    + if r > high_r -> **VH**
    + else if r > 0 -> **H**
	+ else if r is not < 0 -> **M**
    + else -> **L**
* if acceleration, upgrade by 1 step (bound at **VH**)

with by default, *high_r* = 0.05, and at least 3 days of net increase to register as
a sign of acceleration.


```{r }

make_classif <- function(incid, r, r_lower, r_upper, accel,
                         high_r = 0.05,
                         r_limit = 0.02,
                         n_accel = 3,
                         incid_tiers = c(0.4, 0.8)) {

  # The output will be stored as numbers first and converted at the end:
  ## 0 = Minimal
  ## 1 = low
  ## 2 = mid
  ## 3 = high
  ## 4 = very high

  incid_tiers <- sort(incid_tiers)
  
  if (is.na(incid) | is.na(r) | is.na(accel)) return(NA_character_)
  
  ## incid 0-40%
  if (incid < incid_tiers[1]) {
    if (r_lower > high_r) {
      out <- 2
    } else if (r_lower > r_limit) {
      out <- 1
    } else {
      out <- 0
    }
  } else if (incid <= incid_tiers[2]) {
      ## incid 40-80%
      if (r_lower > high_r) {
        out <- 3
      } else if (r_lower > r_limit) {
        out <- 2
      } else if (r_lower > -r_limit) {
        out <- 1
      } else {
        out <- 0
      }
  } else {
    ## incid >80%
    if (r_lower > high_r) {
        out <- 4
      } else if (r_lower > 0) {
        out <- 3
      } else if (r_upper > 0) {
        out <- 2
      } else {
        out <- 1
      }
  } 

  ## add +1 if acceleration
  if (accel >= n_accel & out < 4) {
    out <- out + 1
  }

  c("Minimal", "Low", "Medium", "High", "Very High")[out + 1]
}


## apply classification algo to countries
x <- x %>%
  group_by(iso3) %>%
  mutate(
    classif = make_classif(incid = perc_peak,
                           r = r, r_lower = r_lower, r_upper = r_upper,
                           accel = net_increases),
    classif = factor(classif, levels =  c("Minimal", "Low", "Medium", "High", "Very High"))
  )

## add TPR info
x <- x %>%
  left_join(select(all_elr_extras, -who_region), by = "iso3")

## add classification to x_all
x_all <- x_all %>%
  left_join(x %>% select(iso3, classif), by = "iso3")

```



## Leaderboard table

We reformat the results into `out` and print the table with Low levels and above:

```{r }

out <- x %>%
  arrange(desc(as.integer(classif))) %>%
  select(report_country, iso3,
         classif,
         r, r_lower, r_upper,
         perc_peak, net_increases,
         # removing tpr change for memory reasons
         # tpr_change,
         tpr_recent, tpr_before) %>%
  mutate(r = round(r, 3),
         r_lower = round(r_lower, 3),
         r_upper = round(r_upper, 3),
         perc_peak = round(perc_peak, 3),
         tpr_recent = round(tpr_recent, 3),
         tpr_before = round(tpr_before, 3)
         )

out %>%
  ungroup() %>% 
  mutate(
    perc_peak = paste(round(100 * perc_peak), "%"),
    tpr_recent = if_else(is.na(tpr_recent),
                         "",
                         paste(round(100 * tpr_recent), "%")),
    tpr_before = if_else(is.na(tpr_before),
                         "",
                         paste(round(100 * tpr_before), "%"))) %>%
  rename(Country = report_country,
         "% peak" = perc_peak,
         ELR = classif,
         ) %>%
  select(-iso3) %>% 
  filter(ELR != "Minimal") %>%
  DT::datatable(rownames = FALSE, options = list(pageLength = nrow(.)))

```



## Pinplot

These pinplots shows all countries classified as 'Medium' and above.

### By % incidence peak

```{r fig.height = 7, out.width = "100%"}

scale_signif <- scale_shape_manual(
  "Significant growth/decline",
  values = c(yes = 19, no = 1)
)

scale_asmodee <- scale_colour_steps(
  "Days in last week\nabove\nexpectations",
  low = "#5b7bb5",
  high = "#d10031",
  breaks = 0:7,
  limits = c(0,7))


plot_overall_peaks <- x %>%
  filter(classif %in% c("Medium", "High", "Very High")) %>% 
  ggplot(aes(x = r, y = perc_peak, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_point(aes(shape = r_signif), size = 3) +
  geom_text_repel(
    aes(label = report_country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_y_continuous(labels = scales::percent) +
  scale_signif +
  scale_asmodee +
  labs(x = "Growth rate of daily cases",
       y = "Last week's incidence\n(% of historical peak)") +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2))

plot_overall_peaks

```


### By deaths per capita

```{r  fig.height = 7, out.width = "100%"}

# tadpole plot
plot_overall_deaths <- x %>%
  filter(classif %in% c("Medium", "High", "Very High")) %>% 
  ggplot(aes(x = r, y = death_incid_last_week, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_point(aes(shape = r_signif), size = 3) +
  geom_text_repel(
    aes(label = report_country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_signif +
  scale_asmodee +
  labs(x = "Growth rate of daily cases",
       y = "Weekly COVID-19 deaths per 1M inhabitants") +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2))

plot_overall_deaths

```




## Tadpoles

### By % incidence peak

```{r fig.height = 7, out.width = "100%"}

plot_tadpole_peaks <- x_all %>%
  filter(classif %in% c("Medium", "High", "Very High")) %>% 
  group_by(report_country) %>%
  arrange(last_day) %>% 
  ggplot(aes(x = r, y = perc_peak, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_path(aes(group = report_country, alpha = last_day)) + 
  geom_point(aes(size = last_day, alpha = last_day, shape = r_signif)) +
  geom_text_repel(
    data = filter(x, classif %in% c("Medium", "High", "Very High")),
    aes(label = report_country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_y_continuous(labels = scales::percent) +
  scale_signif +
  scale_asmodee +
  scale_alpha_continuous(trans = "reverse") +
  scale_size_continuous(range = c(0.5, 3), trans = "reverse") +
  labs(x = "Growth rate of daily cases",
       y = "Last week's incidence\n(% of historical peak)") +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2),
    size = "none",
    alpha = "none")

plot_tadpole_peaks


```


### By deaths per capita

```{r  fig.height = 7, out.width = "100%"}

# tadpole plot
plot_tadpole_deaths <- x_all %>%
  filter(classif %in% c("Medium", "High", "Very High")) %>% 
  group_by(report_country) %>%
  arrange(last_day) %>% 
  ggplot(aes(x = r, y = death_incid_last_week, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_path(aes(group = report_country, alpha = last_day)) + 
  geom_point(aes(size = last_day, alpha = last_day, shape = r_signif)) +
  geom_text_repel(
    data = filter(x, classif %in% c("Medium", "High", "Very High")),
    aes(label = report_country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_signif +
  scale_asmodee +
  scale_alpha_continuous(trans = "reverse") +
  scale_size_continuous(range = c(0.5, 3), trans = "reverse") +
  labs(x = "Growth rate of daily cases",
       y = "Weekly COVID-19 deaths per 1M inhabitants") +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2),
    size = "none",
    alpha = "none")

plot_tadpole_deaths

```



## ASMODEE plots

We provide ASMODEE plots for countries classified as "Medium" and above.

```{r }

to_keep <- x %>%
  filter(classif %in% c("Medium", "High", "Very High")) %>%
  pull(report_country)

asmodee_res <- map(all_content, list("results")) %>%
  bind_rows() %>%
  filter(report_country %in% to_keep) %>%
  arrange(report_country)

n_plots <- nrow(asmodee_res)

```

```{r fig.height = 2 * round(n_plots / 3), out.width = "100%"}

plot(asmodee_res, ncol = 3)

```



## VoC / VoI plots

We plot VoC / VoI dynamics over time for the countries classified as 'Medium' and
above, for the last 6 weeks:

```{r }

## re-arrange data into wide format, filter relevant stuff
voi <- dat_voi %>%
  filter(report_country %in% to_keep,
         date >= (Sys.Date() - 7*4),
         date < Sys.Date()) %>%
  arrange(report_country)

## re-order variants
voi <- voi %>%
  mutate(variant = factor(variant),
         variant = forcats::fct_relevel(variant, "Other", after = Inf))

n_plots <- length(unique(voi$report_country))

```

----------------------------

# Export table {.tabset .tabset-fade .tabset-pills}

## Add extra info

We add the following info to the output table:

* Monday of the reporting week
* country info (WHO region, iso3 and population size)

We also renamed columns for compatibility with the decision spreadsheets:

```{r }

# load country info
df_countries <- load_countries_data()

# remove redundant country info in out
out <- out %>%
  ungroup() %>%
  select(-report_country)

# add country info and order table
out_full <- df_countries %>%
  left_join(out, by = "iso3") %>%
  arrange(who_region, report_country)

## add date and rename/reorder columns
out_final <- out_full %>%
  mutate(assessment_week_begins = Sys.Date() %>%
           grates::as_yearweek() %>%
           as.Date()) %>%
  select(
    assessment_date = assessment_week_begins,
    country = report_country,
    iso3, 
    region = who_region,
    population = population,
    growth_rate = r,
    relative_incidence = perc_peak,
    net_increases = net_increases,
    epi_classification = classif,
    growth_rate_lower = r_lower,
    growth_rate_upper = r_upper
  ) %>% 
     mutate(relative_incidence = scales::percent(relative_incidence, accuracy = 0.1))

# join to vaccine coverage/tpr data
elr_extra_to_keep <- all_elr_extras %>%
  select(
    iso3,
    tpr_recent,
    tpr_before,
    tpr_change,
    tpr_last_date_recent,
    tpr_last_date_before,
    vac_oneplus_dose_per_100,
    vac_full_per_100,                      # percent fully vaccinated
    vaccinations_date = as_of_vaccine
         ) %>%
  mutate(
    tpr_recent = scales::percent(tpr_recent, accuracy = 0.1),
    tpr_before = scales::percent(tpr_before, accuracy = 0.1),
    vac_oneplus_dose_per_100 = round(vac_oneplus_dose_per_100, digits = 1),
    vac_full_per_100 = round(vac_full_per_100, digits = 1))

out_final <- out_final %>% 
     left_join(elr_extra_to_keep, by = "iso3")

```



## Export table to xlsx

```{r }

## export asmodee summary
write_excel_csv(out_final, here::here("dynamics_outputs/csv/dynamics_synthesis.csv"), na = "")

## export asmodee summary (dated)
write_excel_csv(out_final, here::here(paste0("dynamics_outputs/csv/dynamics_synthesis_", Sys.Date(), ".csv")), na = "")

```





----------------------------

# Export notes {.tabset .tabset-fade .tabset-pills}

## Compile all notes

We read all the latest notes and compile them into a single character vector:

```{r }

# get file locations
file_paths <- here::here("asmodee_outputs", "notes") %>%
  list.files(pattern = "latest", full.names = TRUE)

# get corresponding regions
file_names <- here::here("asmodee_outputs", "notes") %>%
  list.files(pattern = "latest")
regions <- gsub("[^A-Z]", "", file_names)

# read notes
all_notes <- file_paths %>%
  map(readLines)

# concatenate notes
notes <- lapply(all_notes, function(e) e[1]) %>% unlist()
notes <- paste(regions, "----", notes, "\n", sep = "\n")
notes <- gsub("<br>", "\n", notes)


```


## Export table to xlsx

```{r }

## export notes to file
writeLines(notes, here::here("dynamics_outputs/notes/excluded_countries.txt"))

## export notes to file
writeLines(notes, here::here(paste0("dynamics_outputs/notes/excluded_countries_", Sys.Date(), ".txt")))

```
