---
title: "Using ASMODEE to detect trend changes in COVID-19"
author: "Thibaut Jombart, Henry Laurenson-Schafer, Neale Batra, Finlay Campbell, Yuka Jinnai, Olivier le Polain"
date: '`r format(Sys.time(), "%A %d %B %Y")`'
output: 
  html_document:
    toc: TRUE
    toc_depth: 4
    toc_float: TRUE
    toc_collapse: FALSE
    number_sections: TRUE
    self_contained: TRUE
    highlight: pygments
    theme: spacelab
    code_folding: hide
    <!-- df_print: paged -->
    css: !expr here::here('css', 'style.css')
params:
  who_region: "EURO"
  n_cores: 1
  tadpole_size: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 70,
                      warning = TRUE,
                      message = TRUE)
```

```{r eval = FALSE, echo = FALSE}

rmarkdown::render('regional_analyses.Rmd', params = list(n_cores = 12, who_region = "EURO"))

```


# Prep work {.tabset .tabset-fade .tabset-pills}

## Scripts

```{r warnings = FALSE}

rfextras::load_scripts()

```


## Packages

All CRAN packages can be installed by calling `reportfactory::install_deps()`.
Package installs on remotes are now handled in a separate script
`remote_packages.R` which is loaded alongside other scripts in `scripts/` using:


```{r}

pacman::p_load(
            tidyverse, # data management
            ggrepel, # plots
            future, # parallelisation
            incidence2
)

## remotes loaded separately
library(trending)
library(trendeval)
library(trendbreaker)
library(i2extras)

```


## Parallelisation

We use *future* to handle parallelisation on separate R processes (rather than
multithreading, not supported on windows), which should work across different
platforms:

```{r }

plan("multisession", workers = params$n_cores)

```



## Load the data

```{r }

dat <- load_final_data()
data_stamp <- get_timestamp(dat)
data_stamp
elr_extras <- dat$elr_extras
dat <- dat$trendbreaker
dat
```



## Filter the data

We keep only the countries for the desired WHO region; defaults to EURO:

```{r }

params$who_region
dat <- dat %>%
  filter(who_region %in% params$who_region)

elr_extras <- elr_extras %>% 
  filter(who_region %in% params$who_region)

```





----------------------------

# Trends in cases

## Data

Data will be easier to handle if first converted to an *incidence2* object, as
*asmodee* will recognize these inputs and handle de stratification by country
for us. We also add the *weekday* variable, and retain only the last 6 weeks of
data.

```{r }

cases_i <- dat %>%
  incidence(date,
            counts = c(cases, tests, deaths),
            groups = c(report_country)) %>%
  mutate(weekend = day_of_week(date_index)) %>% # add weekday
  keep_last(7 * 7) %>%
  mutate(weekday = factor(weekdays(date_index)))
  
cases_i

```

We also remove countries which reported less than 100 cases or tests in the last
7 weeks:

```{r }

analysed_countries <- to_keep <- cases_i %>%
  group_by(report_country) %>%
  summarise(total_cases = sum(cases, na.rm = TRUE),
            total_tests = sum(tests, na.rm = TRUE)) %>%
  filter(total_cases > 100 | total_tests > 100) %>%
  pull(report_country)

notes <- NULL
removed_countries <- setdiff(cases_i$report_country, to_keep)
if (length(removed_countries)) {
  msg <- paste0("The following countries reported < 100 cases/tests over the last 7 weeks ",
                "and were not included in further analyses:\n",
                paste(removed_countries, collapse = ", "))
  message(msg)
  notes <- c(notes, gsub("\n", "<br>", msg))
}

n_countries <- length(unique(to_keep))

cases_i <- cases_i %>%
  filter(report_country %in% to_keep)

```

```{r eval = (length(removed_countries) > 0), echo = FALSE, results = "asis"}

cat("**Important note:**<br>")
cat(notes)

```




## Model generators

Here we use some ad-hoc code to handle cases where the training set contains not
one, but 2 temporal trends. The proposed solution is to duplicate each candidate
model of the form `model(response ~ xxx + time)` into:

1. the same model
2. a set of *k* models in the form: `model(response ~ xxx + change_k:time +
time)` where `change_k` is a binary qualitative variable marking before/after
day *k*; the set of models generated will explore various values of *k*


### Adding binary variables to the data

Here we add the *change_k* variables mentioned above; considering we have 6
weeks of data in total, including 5 for training, we allow *k* to vary from day
8 to day 28.

```{r }

cases_i <- cases_i %>%
  mutate(day = as.integer(date_index - min(date_index)))

min_k <- 9
max_k <- 25
k_values <- min_k:max_k
change_k_df <- lapply(k_values,
                       function(k)
                         cases_i %>%
                         group_by(report_country) %>%
                         transmute(if_else(day <= k, "before", "after")) %>%
                         pull(2)) %>%
  data.frame() %>%
  tibble() %>%
  setNames(paste0("change_", k_values))

cases_i <- cases_i %>%
  bind_cols(change_k_df)

```


### Define split models

```{r }

model_grid <- expand.grid(
    c("", "tests"), # with/without tests
    "date_index", # time effect
    c("", "weekend", "weekday"), # optional offsets
    c("", paste("date_index*change", k_values, sep = "_")) # optional split
)

## convert to text
predictors_txt <- model_grid %>%
  apply(1, paste, collapse = " + ")

## cleanup
predictors_txt <- gsub("(\\+[ ]*)+[ ]*\\+", " + ", predictors_txt) # +... + -> +
predictors_txt <- sub("^[ ]*\\+", "", predictors_txt) # heading +
predictors_txt <- sub("\\+[ ]*$", "", predictors_txt) # trailing +
predictors_txt <- sub("^[ ]+", "", predictors_txt) # heading spaces
predictors_txt <- sub("[ ]+$", "", predictors_txt) # trailing spaces
predictors_txt <- sub("[ ]+", " ", predictors_txt) # multiple spaces

## add constant models
predictors_txt <- c("1", predictors_txt)

```

## Candidate models

Candidate models are generated to include:

* Poisson GLM or NegBin GLM
* combinations of:
    + with date effect
    + with / without testing effect
    + with / without weekend effect
    + with / without weekday effect
    + with / without trend shift effect (different slope before/after) with
      changing points on day 8, 9, ... or 28 (first training day is day 0)
	

```{r }

models_txt  <- c(
    ## sprintf("lm_model(cases ~ %s)", predictors_txt), # linear models
    sprintf("glm_model(cases ~ %s, family = gaussian)", predictors_txt) , # Gaussian GLMs
    sprintf("glm_nb_model(cases ~ %s)", predictors_txt) # NegBin GLMs
)

length(models_txt)
head(models_txt)
tail(models_txt)

cases_models <- lapply(models_txt, function(e) eval(parse(text = e)))

```




## Run ASMODEE

Case data can be problematic because of occasional artifactual 'zeros', which
reflect the absence of reporting on a given day, typically followed by a higher
case count reflecting the backlog. Unfortunately, this often throws off the
Negbin GLM, which is most of the time our best option: Poisson GLM drastically
under-estimates variance and sees most points as outliers, and Gaussian GLM can
predict negative numbers.

To circumvent the issue, we use the new feature from *i2extras* called
`flag_low_counts()`, which we use to set to `NA` all counts below a certain
threshold (here, 1/100 of the median). This is done separately for each country.

```{r }

# run asmodee
cases_res <- cases_i %>%
  asmodee(cases_models, k = 7,
          simulate_pi = FALSE)

# keep only successful runs
to_keep <- vapply(cases_res$output, inherits, logical(1), "trendbreaker")
cases_res <- cases_res %>%
  filter(to_keep)

## ## for loop for debugging purposes
## split_x <- cases_i %>%
##   flag_low_counts(set_missing = TRUE, threshold = 0.001) %>%
##   split(cases_i$report_country)
## cases_res <- list()
## for (i in seq_along(split_x)) {
##   cases_res[[i]] <- asmodee(split_x[[i]], cases_models, fixed_k = 7, method = evaluate_aic,
##           include_warnings = FALSE,
##           simulate_pi = FALSE)
## }

```

```{r }

# handle countries with no results
countries_missing_asmodee <- setdiff(analysed_countries, cases_res$report_country)
if (length(countries_missing_asmodee)) {
  msg <- paste0("ASMODEE results could not be produced for the following countries:\n",
                paste(countries_missing_asmodee, collapse = ", "))
  message(msg)
  notes <- c(notes, gsub("\n", "<br>", msg))
}

```

```{r eval = (length(countries_missing_asmodee) > 0), echo = FALSE, results = "asis"}

cat("**Important note:**<br>")
cat(gsub("\n", "<br>", msg))

```


## Plot with all countries

```{r fig.height = 2 * round(n_countries / 3), out.width = "100%"}

cases_res %>%
  plot(ncol = 3)

```



## Countries summaries

We use `summary` to have an overview of outliers by country:

```{r }

cases_outliers <- cases_res %>%
  summary() %>%
  select(Country = report_country,
         n_recent_increases,
         n_recent_decreases,
         p_recent_outliers)

DT::datatable(cases_outliers,
              options = list(
                  pageLength = nrow(cases_outliers)))

```







---------------------------------

# Growth rates

## Growth rates

These estimates of daily growth rates are based on negative binomial GLM fitted
to the last 3 weeks of data. Positive values indicate epidemic growth, and
negative values indicate decline. Confidence intervals including zero indicate
uncertain trends.


```{r fig.height = 8, out.width = "90%"}

#  estimate the growth rates days before (going back: d2 - d10)
## we use future_lapply to speed up computations
df_r  <- future.apply::future_lapply(
  0:params$tadpole_size,
  function(d)
  cases_i %>%
    select(-deaths, -tests) %>% 
    keep_last(3*7 + d) %>%
    keep_first(3*7) %>% 
    fit_curve() %>%
    growth_rate() %>%
    mutate(last_day = d) %>% 
    select(report_country, r, r_lower, r_upper, last_day),
  future.seed = TRUE) %>%
  bind_rows() %>%
  mutate(type = case_when(r_lower > 0 ~ "growth",
                          r_upper < 0 ~ "decline",
                          TRUE ~ "uncertain"),
         type = factor(type, levels = c("growth", "uncertain", "decline")))

```

```{r }

countries_missing_r <- setdiff(analysed_countries, df_r$report_country)

# handle countries with no results
if (length(countries_missing_r)) {
  msg <- paste0("Estimates of *r* could not be produced for the following countries:\n",
                paste(countries_missing_r, collapse = ", "))
  message(msg)
  notes <- c(notes, gsub("\n", "<br>", msg))
}

```

```{r eval = (length(countries_missing_r) > 0), echo = FALSE, results = "asis"}

cat("**Important note:**<br>")
cat(gsub("\n", "<br>", msg))

```



```{r }

scale_type <- scale_color_manual(
  "Incidence trend",
  values = c(growth = "#EC6446",
             decline = "#5DADC6",
             uncertain = "#A9A9A9")
)

df_r %>%
  filter(last_day == 0) %>% 
  ggplot(aes(y = report_country, x = r, color = type)) +
  geom_errorbar(aes(xmin = r_lower, xmax = r_upper)) +
  geom_point() +
  theme_bw() +
  scale_type +
#  scale_x_continuous(labels=scales::percent) +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  labs(x = "Daily growth rate (r)", y = "")

```





--------------------------------

# Overall dynamics

## Merging indicators

```{r }

# Population by country
df_pop <- dat %>%
  group_by(report_country, iso3) %>%
  summarise(population = mean(population, na.rm = TRUE))

# Cumulative incidence over the last week (going back: w1) per 100k inhabitants
df_recent_incid <- future.apply::future_lapply(
   0:params$tadpole_size,
  function(d)
    cases_i %>%
      keep_last(7 + d) %>%
      keep_first(7) %>% 
      group_by(report_country) %>%
      summarise(n_cases = sum(cases),
                n_deaths = sum(deaths)) %>%
      left_join(df_pop) %>%
      mutate(case_incid_last_week = 1e6 * n_cases / population,
             death_incid_last_week = 1e6 * n_deaths / population,
             last_day = d),
  future.seed = TRUE) %>%
  bind_rows()

# Table containing r, incidence per capita, recent increases
df_dynamics <- df_recent_incid %>%
  left_join(df_r) %>%
  left_join(cases_outliers, by = c("report_country" = "Country"))

# Add data on past peak incidence
minimum_peak <- 0

df_peaks <- dat %>%
  filter(who_region == params$who_region,
         date < (data_stamp - 7)) %>%
  group_by(report_country) %>%
  summarise(incid_peak = max(case_new_7d_100k, na.rm = TRUE)) %>%
  mutate(incid_peak = incid_peak * 10) %>% # incidence per 100k -> 1M
  filter(incid_peak >= minimum_peak)

df_dynamics <- df_dynamics %>%
  left_join(df_peaks, by = "report_country") %>%
  mutate(perc_peak = case_incid_last_week / incid_peak)

# Add info on significance of r (> 0 or < 0)
df_dynamics <- df_dynamics %>%
  mutate(r_signif = if_else(r_lower > 0 | r_upper < 0, "yes", "no"))

# Add 'net_increase' over last week
df_dynamics <- df_dynamics %>%
  mutate(net_increases = n_recent_increases - n_recent_decreases,
         net_increases = if_else(net_increases > 0L, net_increases, 0L))

# data with only current time frame
df_dynamics_current <- df_dynamics %>%
  filter(last_day == 0)

```


```{r }

countries_missing_peak <- setdiff(analysed_countries, df_peaks$report_country)

# handle countries with no results
if (length(countries_missing_peak)) {
  msg <- sprintf(
    "Past incidence peaks could not be found (no case previously reported) for the following countries:\n%s",
    minimum_peak,
    paste(countries_missing_peak, collapse = ", "))
  message(msg)
  notes <- c(notes, gsub("\n", "<br>", msg))
}

```

```{r eval = (length(countries_missing_peak) > 0), echo = FALSE, results = "asis"}

cat("**Important note:**<br>")
cat(gsub("\n", "<br>", msg))

```



## Deaths, growth rates and changes

### Scales

```{r }

scale_signif <- scale_shape_manual(
  "Significant growth/decline",
  values = c(yes = 19, no = 1)
)

scale_asmodee <- scale_colour_steps(
  "Days in last week\nabove\nexpectations",
  low = "#5b7bb5",
  high = "#d10031",
  breaks = 0:7,
  limits = c(0,7))

```


### Main figure

This figure reports:

* growth rate based on the last 3 weeks of data
* last weeks's deaths per capita
* significant increases detected by ASMODEE in the last week

```{r  fig.height = 7, out.width = "100%"}

# tadpole plot
plot_overall_deaths <- df_dynamics_current %>%
  ggplot(aes(x = r, y = death_incid_last_week, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_point(aes(shape = r_signif), size = 3) +
  geom_text_repel(
    data = df_dynamics_current,
    aes(label = report_country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_signif +
  scale_asmodee +
  labs(x = "Growth rate of daily cases",
       y = "Weekly COVID-19 deaths per 1M inhabitants",
       subtitle = sprintf("Last data update: %s",
                          format(data_stamp, "%A %d %B %Y"))) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2))

plot_overall_deaths



```

### Tadpole version

```{r  fig.height = 7, out.width = "100%"}

# tadpole plot
plot_tadpole_deaths <- df_dynamics %>%
  group_by(report_country) %>%
  arrange(last_day) %>% 
  ggplot(aes(x = r, y = death_incid_last_week, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_path(aes(group = report_country, alpha = last_day)) + 
  geom_point(aes(size = last_day, alpha = last_day, shape = r_signif)) +
  geom_text_repel(
    data = df_dynamics_current,
    aes(label = report_country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_signif +
  scale_asmodee +
  scale_alpha_continuous(trans = "reverse") +
  scale_size_continuous(range = c(0.5, 3), trans = "reverse") +
  labs(x = "Growth rate of daily cases",
       y = "Weekly COVID-19 deaths per 1M inhabitants",
       subtitle = sprintf("Last data update: %s",
                          format(data_stamp, "%A %d %B %Y"))) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2),
    size = "none",
    alpha = "none")

plot_tadpole_deaths

```


## Relative cases, growth and changes

### Main figure

This figure reports:

* growth rate based on the last 3 weeks of data
* last weeks's case incidence per capita, expressed as a proportion of the past
  peak incidence
* significant increases detected by ASMODEE in the last week

```{r  fig.height = 7, out.width = "100%"}

plot_overall_peaks <- df_dynamics_current %>%
  ggplot(aes(x = r, y = perc_peak, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_point(aes(shape = r_signif), size = 3) +
  geom_text_repel(
    data = df_dynamics_current,
    aes(label = report_country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_y_continuous(labels = scales::percent) +
  scale_signif +
  scale_asmodee +
  labs(x = "Growth rate of daily cases",
       y = "Last week's incidence\n(% of historical peak)",
       subtitle = sprintf("Last data update: %s",
                          format(data_stamp, "%A %d %B %Y"))) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2))

plot_overall_peaks

```


### Tadpole version

```{r  fig.height = 7, out.width = "100%"}

plot_tadpole_peaks <- df_dynamics %>%
  group_by(report_country) %>%
  arrange(last_day) %>% 
  ggplot(aes(x = r, y = perc_peak, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_path(aes(group = report_country, alpha = last_day)) + 
  geom_point(aes(size = last_day, alpha = last_day, shape = r_signif)) +
  geom_text_repel(
    data = df_dynamics_current,
    aes(label = report_country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_y_continuous(labels = scales::percent) +
  scale_signif +
  scale_asmodee +
  scale_alpha_continuous(trans = "reverse") +
  scale_size_continuous(range = c(0.5, 3), trans = "reverse") +
  labs(x = "Growth rate of daily cases",
       y = "Last week's incidence\n(% of historical peak)",
       subtitle = sprintf("Last data update: %s",
                          format(data_stamp, "%A %d %B %Y"))) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2),
    size = "none",
    alpha = "none")

plot_tadpole_peaks

```





---------------------------

# Test positivity rates (TPR)

## Changes in TPR

The changes in TPR is determined by comparing the last estimated weekly TPR
(i.e. including the last available data, excluding artifactual trailing zeros)
to the TPR of the week before. TPR estimates as well as the corresponding raw
numbers are generated by `assemble_data` and stored in `elr_extras`, and are
identical to the values reported in the 
[shiny app](https://worldhealthorg.shinyapps.io/covid/). 
We assess the significance of changes by comparing the respective Binomial 95%
confidence intervals:


```{r, eval= FALSE}

elr_extras <- elr_extras %>%
  mutate(
    tpr_change = case_when(
      is.na(tpr_recent) | is.na(tpr_before) ~ NA_character_,
      tpr_lower_recent >  tpr_upper_before ~ "increasing",
      tpr_upper_recent <  tpr_lower_before ~ "decreasing",
      TRUE ~ "constant"),
    tpr_change = factor(tpr_change,
                        levels = c("decreasing", "constant", "increasing")))


```





------------------------------

# Export results

## Create needed directories

We create output directories if they don't already exist:

```{r }

target <- here::here("asmodee_outputs")
if (!dir.exists(target)) dir.create(target)

targets <- c("rds", "png", "notes")
for (e in targets) {
  target <- here::here("asmodee_outputs", e)
  if (!dir.exists(target)) dir.create(target)
}

```


## Save ASMODEE outputs 

We export the following products for further uses in other systems, in an object
called `asmodee_outputs`; all objects are stored in */asmodee_outputs/rds/*:

* `summary`: summary of the ASMODEE results
* `results`: outputs of ASMODEE
* `plot_overall_deaths`: *ggplot2* object of the overall dynamics plot using
  death per capita on the y-axis
* `plot_overall_peaks`: *ggplot2* object of the overall dynamics plot using
  incidence as percentage of historical peak on the y-axis
* `df_dynamics`: a `data.frame` containing all the required info to recreate
  either global dynamics plots
* `elr_extras`: a `data.frame` containing additional information for countries,
  including TPR and vaccination coverage
* `timestamp`: the timestamp corresponding to the date of the dataset used in
  the analyses

```{r }

asmodee_outputs <- list(
  summary = cases_outliers,
  results = cases_res,
  plot_overall_deaths = plot_overall_deaths,
  plot_overall_peaks = plot_overall_peaks,
  df_dynamics = df_dynamics,
  elr_extras = elr_extras,
  timestamp = data_stamp
)

## export asmodee summary
file_name <- sprintf("asmodee_outputs_%s_%s.rds",
                     paste(params$who_region, collapse = "_"),
                     data_stamp)
path_out <- here::here("asmodee_outputs", "rds", file_name)
rio::export(asmodee_outputs, path_out)

## make a copy as latest file
latest_file_name <- sprintf("asmodee_outputs_%s_latest.rds",
                            paste(params$who_region, collapse = "_")
                            )
latest_path_out <- here::here("asmodee_outputs", "rds", latest_file_name)
file.copy(from = path_out, to = latest_path_out, overwrite = TRUE)

  
```



## Export main figures

Here we save the two key figures showing 'overall dynamics' to *png* files in
*asmodee_outputs/png/*:

```{r fig.show='hide'}

# Pinplots
## Deaths PC as y axis
fig_name <- sprintf("pinplot_deaths_%s_%s.png",
                    paste(params$who_region, collapse = "_"),
                    data_stamp)
fig_path <- here::here("asmodee_outputs", "png", fig_name)

png(fig_path, width = 22, height = 18, units = "cm", res = 150)
print(plot_overall_deaths)
dev.off()

## Make a copy as 'latest'
latest_fig_name <- sprintf("pinplot_deaths_%s_latest.png",
                           paste(params$who_region, collapse = "_")
                           )
latest_fig_path <- here::here("asmodee_outputs", "png", latest_fig_name)
file.copy(from = fig_path, to = latest_fig_path, overwrite = TRUE)


## cases as % historical peak as y axis
fig_name <- sprintf("pinplot_peaks_%s_%s.png",
                    paste(params$who_region, collapse = "_"),
                    data_stamp)
fig_path <- here::here("asmodee_outputs", "png", fig_name)

png(fig_path, width = 22, height = 18, units = "cm", res = 150)
print(plot_overall_peaks)
dev.off()

## Make a copy as 'latest'
latest_fig_name <- sprintf("pinplot_peaks_%s_latest.png",
                           paste(params$who_region, collapse = "_")
                           )
latest_fig_path <- here::here("asmodee_outputs", "png", latest_fig_name)
file.copy(from = fig_path, to = latest_fig_path, overwrite = TRUE)


# Tadpoles
## Deaths PC as y axis
fig_name <- sprintf("tadpole_deaths_%s_%s.png",
                    paste(params$who_region, collapse = "_"),
                    data_stamp)
fig_path <- here::here("asmodee_outputs", "png", fig_name)

png(fig_path, width = 22, height = 18, units = "cm", res = 150)
print(plot_tadpole_deaths)
dev.off()

# Make a copy as 'latest'
latest_fig_name <- sprintf("tadpole_deaths_%s_latest.png",
                           paste(params$who_region, collapse = "_")
                           )
latest_fig_path <- here::here("asmodee_outputs", "png", latest_fig_name)
file.copy(from = fig_path, to = latest_fig_path, overwrite = TRUE)


## cases as % historical peak as y axis
fig_name <- sprintf("tadpole_peaks_%s_%s.png",
                    paste(params$who_region, collapse = "_"),
                    data_stamp)
fig_path <- here::here("asmodee_outputs", "png", fig_name)

png(fig_path, width = 22, height = 18, units = "cm", res = 150)
print(plot_tadpole_peaks)
dev.off()

## Make a copy as 'latest'
latest_fig_name <- sprintf("tadpole_peaks_%s_latest.png",
                           paste(params$who_region, collapse = "_")
                           )
latest_fig_path <- here::here("asmodee_outputs", "png", latest_fig_name)
file.copy(from = fig_path, to = latest_fig_path, overwrite = TRUE)

```



## Export notes

```{r }

if (is.null(notes)) {
  notes <- "Nothing to report. All countries were successfully included in the analyses."
}

## export asmodee summary
file_name <- sprintf("analysis_notes_%s_%s.md",
                     paste(params$who_region, collapse = "_"),
                     data_stamp)
path_out <- here::here("asmodee_outputs", "notes", file_name)
writeLines(notes, path_out, sep = "\n<br>\n")

## make a copy as latest file
latest_file_name <- sprintf("analysis_notes_%s_latest.md",
                            paste(params$who_region, collapse = "_")
                            )
latest_path_out <- here::here("asmodee_outputs", "notes", latest_file_name)
file.copy(from = path_out, to = latest_path_out, overwrite = TRUE)

```
