---
title: "Listing situations of concern"
author: "Thibaut Jombart"
date: '`r format(Sys.time(), "%A %d %B %Y")`'
output: 
  html_document:
    toc: TRUE
    toc_depth: 4
    toc_float: TRUE
    toc_collapse: FALSE
    number_sections: TRUE
    highlight: pygments
    theme: spacelab
    code_folding: hide
    <!-- df_print: paged -->
    css: !expr here::here('css', 'style.css')
params:
  who_region: "EURO"
  n_cores: 1
  tadpole_size: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 70,
                      warning = TRUE,
                      message = TRUE)
```

```{r eval = FALSE, echo = FALSE}

rmarkdown::render('trendbreaker.Rmd', params = list(n_cores = 12, who_region = "EURO"))

```


# Prep work {.tabset .tabset-fade .tabset-pills}

## Scripts

```{r warnings = FALSE}

rfextras::load_scripts()

```


## Packages

All CRAN packages can be installed by calling `reportfactory::install_deps()`.
Package installs on remotes are now handled in a separate script
`remote_packages.R` which is loaded alongside other scripts in `scripts/` using:


```{r}

pacman::p_load(
  tidyverse,
  rio
)

```



## Load the data

We load the `$df_dynamics` components of the list of results stored for each WHO
region, output by `trendbreaker.Rmd`:

```{r }

all_files <- list.files(here::here("asmodee_outputs"),
                        pattern = "latest.rds",
                        recursive = TRUE,
                        full.names = TRUE)

all_content <- map(all_files, rio::import)
regions <- all_files %>%
  gsub("[[:lower:][:punct:]]", "", .)

all_df <- all_content %>%
  map(list("df_dynamics")) %>%
  setNames(regions)

```

Add region data and assemble into a single `data.frame`, keeping only the
current day:

```{r }


for (i in seq_along(regions)) {
  all_df[[i]] <- mutate(all_df[[i]], who_region = regions[i])
}

x <- all_df %>%
  bind_rows() %>%
  filter(last_day == 0)

```




# Generate pre-classification

## Classification algorithm

We define the following classes for levels of concerns:

* not listed
* *low*: average incidence *and* no decline
* *mid*: high incidence *or* fast growth, **or** average incidence *and* growth
* *high*: high/very high incidence *and* growing, **or** low incidence *and* growing fast
  very fast
* *very high*: high/very high incidence *and* growing fast/very fast

We use the following classification rules, depending on the values of the growth
rate (*r*), current incidence (as % of previous peak) and signs of acceleration
(ASMODEE net increases over last week):

* if incid < 50%:
    + if r > high_r -> **M**
	+ else -> NL
* if 50% < incid  100%:
    + if r > high_r -> **H**
    + else if r > 0.02 -> **M**
	+ else if r is not < 0 -> **L**
    + else -> **NL**
* if incid > 100%:
    + if r > high_r -> **VH**
    + else if r > 0 -> **H**
	+ else if r is not < 0 -> **M**
    + else -> **NL**
* if acceleration, upgrade by 1 step (bound at **VH**)

with by default, *high_r* = 0.05, and at least 3 days of net increase to register as
a sign of acceleration.


```{r }

make_classif <- function(incid, r, r_lower, r_upper, accel,
                         high_r = 0.05,
                         r_limit = 0.02,
                         n_accel = 3) {

  # The output will be stored as numbers first and converted at the end:
  ## 0 = NL
  ## 1 = low
  ## 2 = mid
  ## 3 = high
  ## 4 = very high
  
  if (incid < 0.5) {
    if (r > high_r) {
      out <- 2
    } else {
      out <- 0
    }
  } else if (incid < 1) {
      if (r > high_r) {
        out <- 3
      } else if (r > r_limit) {
        out <- 2
      } else if (r)
  } else {

  } 

}



## x %>%
##   group_by(report_country) %>%
##   mutate(
##     classif = case_when(
      
##     )
##       )

```


## Export notes

```{r eval = FALSE}

if (is.null(notes)) {
  notes <- "Nothing to report. All countries were successfully included in the analyses."
}

## export asmodee summary
file_name <- sprintf("analysis_notes_%s_%s.md",
                     paste(params$who_region, collapse = "_"),
                     data_stamp)
path_out <- here::here("asmodee_outputs", "notes", file_name)
writeLines(notes, path_out, sep = "\n<br>\n")

## make a copy as latest file
latest_file_name <- sprintf("analysis_notes_%s_latest.md",
                            paste(params$who_region, collapse = "_")
                            )
latest_path_out <- here::here("asmodee_outputs", "notes", latest_file_name)
file.copy(from = path_out, to = latest_path_out, overwrite = TRUE)

```
