---
title: "Using ASMODEE to detect trend changes in COVID-19"
author: "Thibaut Jombart, Neale Batra, Finlay Campbell, Yuka Jinnai, Henry Laurenson-Schafer, Olivier le Polain"
date: '`r format(Sys.time(), "%A %d %B %Y")`'
output: 
  html_document:
    toc: TRUE
    toc_depth: 4
    toc_float: TRUE
    toc_collapse: FALSE
    number_sections: TRUE
    highlight: pygments
    theme: spacelab
    code_folding: hide
    <!-- df_print: paged -->
    css: !expr here::here('css', 'style.css')
params:
  who_region: "EURO"
  n_cores: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 70,
                      warning = TRUE,
                      message = TRUE)
```


# Prep work {.tabset .tabset-fade .tabset-pills}

## Scripts

```{r warnings = FALSE}

rfextras::load_scripts()

```


## Packages

All CRAN packages can be installed by calling `reportfactory::install_deps()`.
Package installs on remotes are now handled in a separate script
`remote_packages.R` which is loaded alongside other scripts in `scripts/` using:


```{r}

pacman::p_load(
            tidyverse, # data management
            ggrepel, # plots
            future # parallelisation
        )

## remotes loaded separately
library(trending)
library(trendeval)
library(trendbreaker)
library(incidence2)
library(i2extras)

```


## Parallelisation

We use *future* to handle parallelisation on separate R processes (rather than
multithreading, not supported on windows), which should work across different
platforms:

```{r }

plan("multisession", workers = params$n_cores)

```



## Load the data

```{r }

dat <- load_final_data()
data_stamp <- get_timestamp(dat)
data_stamp
dat

```



## Filter the data

We keep only the countries for the desired WHO region; defaults to EURO:

```{r }

params$who_region
dat <- dat %>%
  filter(who_region %in% params$who_region)

```





----------------------------

# Trends in cases

## Data

Data will be easier to handle if first converted to an *incidence2* object, as
*asmodee* will recognize these inputs and handle de stratification by country
for us. We also add the *weekday* variable, and retain only the last 6 weeks of
data.

```{r }

cases_i <- dat %>%
  incidence(date, counts = c(cases, tests, deaths), groups = report_country) %>%
  mutate(weekend = day_of_week(date_index)) %>% # add weekday
  keep_last(7 * 7) %>%
  mutate(weekday = factor(weekdays(date_index)))
  
cases_i

```

We also remove countries which reported less than 100 cases or tests in the last
7 weeks:

```{r }

analysed_countries <- to_keep <- cases_i %>%
  group_by(report_country) %>%
  summarise(total_cases = sum(cases, na.rm = TRUE),
            total_tests = sum(tests, na.rm = TRUE)) %>%
  filter(total_cases > 100 | total_tests > 100) %>%
  pull(report_country)

notes <- NULL
removed_countries <- setdiff(cases_i$report_country, to_keep)
if (length(removed_countries)) {
  msg <- paste0("The following countries reported < 100 cases/tests over the last 7 weeks ",
                "and were not included in further analyses:\n",
                paste(removed_countries, collapse = ", "))
  message(msg)
  notes <- c(notes, gsub("\n", "<br>", msg))
}

n_countries <- length(unique(to_keep))

cases_i <- cases_i %>%
  filter(report_country %in% to_keep)

```

```{r eval = (length(removed_countries) > 0), echo = FALSE, results = "asis"}

cat("**Important note:**<br>")
cat(notes)

```




## Model generators

Here we use some ad-hoc code to handle cases where the training set contains not
one, but 2 temporal trends. The proposed solution is to duplicate each candidate
model of the form `model(response ~ xxx + time)` into:

1. the same model
2. a set of *k* models in the form: `model(response ~ xxx + change_k:time +
time)` where `change_k` is a binary qualitative variable marking before/after
day *k*; the set of models generated will explore various values of *k*


### Adding binary variables to the data

Here we add the *change_k* variables mentioned above; considering we have 6
weeks of data in total, including 5 for training, we allow *k* to vary from day
8 to day 28.

```{r }

cases_i <- cases_i %>%
  mutate(day = as.integer(date_index - min(date_index)))

min_k <- 9
max_k <- 25
k_values <- min_k:max_k
change_k_df <- lapply(k_values,
                       function(k)
                         cases_i %>%
                         group_by(report_country) %>%
                         transmute(if_else(day <= k, "before", "after")) %>%
                         pull(2)) %>%
  data.frame() %>%
  tibble() %>%
  setNames(paste0("change_", k_values))

cases_i <- cases_i %>%
  bind_cols(change_k_df)

```


### Define split models

```{r }

model_grid <- expand.grid(
    c("", "tests"), # with/without tests
    "date_index", # time effect
    c("", "weekend", "weekday"), # optional offsets
    c("", paste("date_index*change", k_values, sep = "_")) # optional split
)

## convert to text
predictors_txt <- model_grid %>%
  apply(1, paste, collapse = " + ")

## cleanup
predictors_txt <- gsub("(\\+[ ]*)+[ ]*\\+", " + ", predictors_txt) # +... + -> +
predictors_txt <- sub("^[ ]*\\+", "", predictors_txt) # heading +
predictors_txt <- sub("\\+[ ]*$", "", predictors_txt) # trailing +
predictors_txt <- sub("^[ ]+", "", predictors_txt) # heading spaces
predictors_txt <- sub("[ ]+$", "", predictors_txt) # trailing spaces
predictors_txt <- sub("[ ]+", " ", predictors_txt) # multiple spaces

## add constant models
predictors_txt <- c("1", predictors_txt)

```

## Candidate models

Candidate models are generated to include:

* Poisson GLM or NegBin GLM
* combinations of:
    + with date effect
    + with / without testing effect
    + with / without weekend effect
    + with / without weekday effect
    + with / without trend shift effect (different slope before/after) with
      changing points on day 8, 9, ... or 28 (first training day is day 0)
	

```{r }

models_txt  <- c(
    ## sprintf("lm_model(cases ~ %s)", predictors_txt), # linear models
    sprintf("glm_model(cases ~ %s, family = gaussian)", predictors_txt) , # Gaussian GLMs
    sprintf("glm_nb_model(cases ~ %s)", predictors_txt) # NegBin GLMs
)

length(models_txt)
head(models_txt)
tail(models_txt)

cases_models <- lapply(models_txt, function(e) eval(parse(text = e)))

```




## Run ASMODEE

Case data can be problematic because of occasional artifactual 'zeros', which
reflect the absence of reporting on a given day, typically followed by a higher
case count reflecting the backlog. Unfortunately, this often throws off the
Negbin GLM, which is most of the time our best option: Poisson GLM drastically
under-estimates variance and sees most points as outliers, and Gaussian GLM can
predict negative numbers.

To circumvent the issue, we use the new feature from *i2extras* called
`flag_low_counts()`, which we use to set to `NA` all counts below a certain
threshold (here, 1/100 of the median). This is done separately for each country.

```{r }

# run asmodee
cases_res <- cases_i %>%
##  flag_low_counts(set_missing = TRUE, threshold = 0.001) %>%
  asmodee(cases_models, k = 7,
          simulate_pi = FALSE)


# keep only successful runs
to_keep <- vapply(cases_res$output, inherits, logical(1), "trendbreaker")
cases_res <- cases_res %>%
  filter(to_keep)

## ## for loop for debugging purposes
## split_x <- cases_i %>%
##   flag_low_counts(set_missing = TRUE, threshold = 0.001) %>%
##   split(cases_i$report_country)
## cases_res <- list()
## for (i in seq_along(split_x)) {
##   cases_res[[i]] <- asmodee(split_x[[i]], cases_models, fixed_k = 7, method = evaluate_aic,
##           include_warnings = FALSE,
##           simulate_pi = FALSE)
## }

```

```{r }

# handle countries with no results
countries_missing_asmodee <- setdiff(analysed_countries, cases_res$report_country)
if (length(countries_missing_asmodee)) {
  msg <- paste0("ASMODEE results could not be produced for the following countries:\n",
                paste(countries_missing_asmodee, collapse = ", "))
  message(msg)
  notes <- c(notes, gsub("\n", "<br>", msg))
}

```

```{r eval = (length(countries_missing_asmodee) > 0), echo = FALSE, results = "asis"}

cat("**Important note:**<br>")
cat(gsub("\n", "<br>", msg))

```


## Plot with all countries

```{r fig.height = 2 * round(n_countries / 3), out.width = "100%"}

cases_res %>%
  plot(ncol = 3)

```



## Countries summaries

We use `summary` to have an overview of outliers by country:

```{r }

cases_outliers <- cases_res %>%
  summary() %>%
  select(Country = report_country,
         n_recent_increases,
         n_recent_decreases,
         p_recent_outliers)

DT::datatable(cases_outliers,
              options = list(
                  pageLength = nrow(cases_outliers)))

```







---------------------------------

# Growth rates

## Growth rates

These estimates of daily growth rates are based on negative binomial GLM fitted
to the last 3 weeks of data. Positive values indicate epidemic growth, and
negative values indicate decline. Confidence intervals including zero indicate
uncertain trends.


```{r fig.height = 8, out.width = "90%"}

# estimate the growth rates on last 3 weeks (going back: w1 - w3)
r_df <- cases_i %>%
  select(-deaths, -tests) %>% 
  keep_last(3*7) %>%
  fit_curve() %>%
  growth_rate() %>%
  mutate(type = case_when(r_lower > 0 ~ "growth",
                          r_upper < 0 ~ "decline",
                          TRUE ~ "uncertain"),
         factor(type, levels = c("growth", "uncertain", "decline")))

#  estimate the growth rates weeks before (going back: w2 - w4)
r_df_w2_w4 <- cases_i %>%
  select(-deaths, -tests) %>% 
  keep_last(4*7) %>%
  keep_first(3*7) %>% 
  fit_curve() %>%
  growth_rate() %>%
  rename(r_2_4 = r) %>% 
  select(report_country, r_2_4)

#  estimate the growth rates weeks before (going back: w3 - w5)
r_df_w3_w5 <- cases_i %>%
  select(-deaths, -tests) %>% 
  keep_last(5*7) %>%
  keep_first(3*7) %>% 
  fit_curve() %>%
  growth_rate() %>%
  rename(r_3_5 = r) %>%
  select(report_country, r_3_5)


# merge estimates
r_df <- r_df %>%
  left_join(r_df_w2_w4) %>%
  left_join(r_df_w3_w5)

```

```{r }

countries_missing_r <- setdiff(analysed_countries, r_df$report_country)

# handle countries with no results
if (length(countries_missing_r)) {
  msg <- paste0("Estimates of *r* could not be produced for the following countries:\n",
                paste(countries_missing_r, collapse = ", "))
  message(msg)
  notes <- c(notes, gsub("\n", "<br>", msg))
}

```

```{r eval = (length(countries_missing_r) > 0), echo = FALSE, results = "asis"}

cat("**Important note:**<br>")
cat(gsub("\n", "<br>", msg))

```



```{r }

scale_type <- scale_color_manual(
  "Incidence trend",
  values = c(growth = "#EC6446",
             decline = "#5DADC6",
             uncertain = "#A9A9A9")
)

r_df %>%
  ggplot(aes(y = report_country, x = r, color = type)) +
  geom_errorbar(aes(xmin = r_lower, xmax = r_upper)) +
  geom_point() +
  theme_bw() +
  scale_type +
#  scale_x_continuous(labels=scales::percent) +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  labs(x = "Daily growth rate (r)", y = "")

```





--------------------------------

# Overall dynamics

## Merging indicators

```{r }

# Population by country
df_pop <- dat %>%
  group_by(report_country) %>%
  summarise(population = mean(population, na.rm = TRUE))

# Cumulative incidence over the last week (going back: w1) per 100k inhabitants
df_recent_incid <- cases_i %>%
  keep_last(7) %>%
  group_by(report_country) %>%
  summarise(n_cases = sum(cases),
            n_deaths = sum(deaths)) %>%
  left_join(df_pop) %>%
  mutate(case_incid_last_week = 1e6 * n_cases / population,
         death_incid_last_week = 1e6 * n_deaths / population)

# Cumulative incidence over the previous week (going back: w2) per 100k inhabitants
df_recent_incid_w2 <- cases_i %>%
  keep_last(2*7) %>%
  keep_first(7) %>%
  group_by(report_country) %>%
  summarise(n_cases = sum(cases),
            n_deaths = sum(deaths)) %>%
  left_join(df_pop) %>%
  mutate(case_incid_week_2 = 1e6 * n_cases / population,
         death_incid_week_2 = 1e6 * n_deaths / population) %>%
  select(report_country, case_incid_week_2, death_incid_week_2)

# Cumulative incidence over the previous week (going back: w3) per 100k inhabitants
df_recent_incid_w3 <- cases_i %>%
  keep_last(3*7) %>%
  keep_first(7) %>%
  group_by(report_country) %>%
  summarise(n_cases = sum(cases),
            n_deaths = sum(deaths)) %>%
  left_join(df_pop) %>%
  mutate(case_incid_week_3 = 1e6 * n_cases / population,
         death_incid_week_3 = 1e6 * n_deaths / population) %>%
  select(report_country, case_incid_week_3, death_incid_week_3)

# Combine cumulative incidences
df_recent_incid <- df_recent_incid %>%
  left_join(df_recent_incid_w2) %>%
  left_join(df_recent_incid_w3)


# Table containing r, recent increases, incidence per 1000
df_dynamics <- cases_outliers %>%
  select(Country, n_recent_increases, n_recent_decreases) %>%
  right_join(select(r_df,
                    Country = report_country,
                    r, r_lower, r_upper,
                    r_2_4, r_3_5)) %>%
  left_join(df_recent_incid, by = c("Country" = "report_country"))

# Add data on past peak incidence, excluding the last 90 days of data
minimum_peak <- 15

df_peaks <- dat %>%
  filter(who_region == params$who_region,
         date <= (data_stamp - 90)) %>%
  group_by(report_country) %>%
  summarise(incid_peak = max(case_new_7d_100k, na.rm = TRUE)) %>%
  mutate(incid_peak = incid_peak * 10) %>% # incidence per 100k -> 1M
  filter(incid_peak >= minimum_peak)

df_dynamics <- df_dynamics %>%
  left_join(df_peaks %>% rename(Country = report_country)) %>%
  mutate(perc_peak = case_incid_last_week / incid_peak,
         perc_peak_2 = case_incid_week_2 / incid_peak,
         perc_peak_3 = case_incid_week_3 / incid_peak)

# Add info on significance of r (> 0 or < 0)
df_dynamics <- df_dynamics %>%
  mutate(r_signif = if_else(r_lower > 0 | r_upper < 0, "yes", "no"))

# Add 'net_increase' over last week
df_dynamics <- df_dynamics %>%
  mutate(net_increases = n_recent_increases - n_recent_decreases,
         net_increases = if_else(net_increases > 0L, net_increases, 0L))

```


```{r }

countries_missing_peak <- setdiff(analysed_countries, df_peaks$report_country)

# handle countries with no results
if (length(countries_missing_peak)) {
  msg <- sprintf(
    "Past incidence peaks were very low (<%d/M/week) and ignored for the following countries:\n%s",
    minimum_peak,
    paste(countries_missing_peak, collapse = ", "))
  message(msg)
  notes <- c(notes, gsub("\n", "<br>", msg))
}

```

```{r eval = (length(countries_missing_peak) > 0), echo = FALSE, results = "asis"}

cat("**Important note:**<br>")
cat(gsub("\n", "<br>", msg))

```



## Deaths, growth rates and changes

This figure reports:

* growth rate based on the last 3 weeks of data
* last weeks's deaths per capita
* significant increases detected by ASMODEE in the last week

```{r  fig.height = 7, out.width = "100%"}

scale_signif <- scale_shape_manual(
  "Significant growth/decline",
  values = c(yes = 19, no = 1)
)


scale_asmodee <- scale_colour_steps(
  "Days in last week\nabove\nexpectations",
  low = "#5b7bb5",
  high = "#d10031",
  breaks = 0:7,
  limits = c(0,7))

# version 1
plot_overall_deaths_1 <- df_dynamics %>%
  ggplot(aes(x = r, y = death_incid_last_week, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_point(aes(shape = r_signif, x = r_3_5, y = death_incid_week_3),
             color = grey(0.2), alpha = 0.2, size = 1) +
  geom_segment(aes(x = r_2_4, y = death_incid_week_2,
                   xend = r_3_5, yend = death_incid_week_3),
               color = grey(0.2), alpha = 0.2, size = 1) +
  geom_point(aes(shape = r_signif, x = r_2_4, y = death_incid_week_2),
             color = grey(0.2), alpha = 0.5, size = 2) +
  geom_segment(aes(xend = r_2_4, yend = death_incid_week_2),
               color = grey(0.2), alpha = 0.5, size = 2) +
  geom_point(aes(shape = r_signif), size = 3) +
  geom_text_repel(aes(label = Country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_signif +
  scale_asmodee +
  labs(x = "Growth rate of daily cases",
       y = "Weekly COVID-19 deaths per 1M inhabitants",
       subtitle = sprintf("Last data update: %s",
                          format(data_stamp, "%A %d %B %Y"))) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2))

plot_overall_deaths_1



# version 2
plot_overall_deaths_2 <- df_dynamics %>%
  ggplot(aes(x = r, y = death_incid_last_week, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_point(aes(shape = r_signif, x = r_3_5, y = death_incid_week_3),
             color = grey(0.2), alpha = 0.2, size = 1) +
  geom_segment(aes(x = r_3_5, y = death_incid_week_3,
                   xend = r_2_4, yend = death_incid_week_2),
               color = grey(0.2), alpha = 0.2,
               arrow = arrow(length = unit(0.02, "npc"), angle = 15)) +
  geom_point(aes(shape = r_signif, x = r_2_4, y = death_incid_week_2),
             color = grey(0.2), alpha = 0.5, size = 2) +
  geom_segment(aes(x = r_2_4, y = death_incid_week_2,
                   xend = r, yend = death_incid_last_week),
               color = grey(0.2), alpha = 0.5,
               arrow = arrow(length = unit(0.02, "npc"), angle = 15)) +
  geom_point(aes(shape = r_signif), size = 3) +
  geom_text_repel(aes(label = Country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_signif +
  scale_asmodee +
  labs(x = "Growth rate of daily cases",
       y = "Weekly COVID-19 deaths per 1M inhabitants",
       subtitle = sprintf("Last data update: %s",
                          format(data_stamp, "%A %d %B %Y"))) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2))

plot_overall_deaths_2



```


## Relative cases, growth and changes


This figure reports:

* growth rate based on the last 3 weeks of data
* last weeks's case incidence per capita, expressed as a proportion of the past
  peak incidence
* significant increases detected by ASMODEE in the last week

```{r  fig.height = 7, out.width = "100%"}

plot_overall_peaks <- df_dynamics %>%
  ggplot(aes(x = r, y = perc_peak, color = net_increases)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2, color = "salmon") +
  geom_point(aes(shape = r_signif), size = 3) +
  geom_text_repel(aes(label = Country), max.overlaps = 100,
                  max.time = 2, max.iter = 1e5) +
  scale_y_continuous(labels = scales::percent) +
  scale_signif +
  scale_asmodee +
  labs(x = "Growth rate of daily cases",
       y = "Last week's incidence\n(% of historical peak, excluding last 90 days)",
       subtitle = sprintf("Last data update: %s",
                          format(data_stamp, "%A %d %B %Y"))) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.title.align = 0.5) +
  guides(
    color = guide_colorsteps(order = 1),
    shape = guide_legend(title.position = "top", order = 2))

plot_overall_peaks

```




------------------------------

# Export results

## Create needed directories

We create output directories if they don't already exist:

```{r }

target <- here::here("asmodee_outputs")
if (!dir.exists(target)) dir.create(target)

targets <- c("rds", "png", "notes")
for (e in targets) {
  target <- here::here("asmodee_outputs", e)
  if (!dir.exists(target)) dir.create(target)
}

```


## Save ASMODEE outputs 

We export the following products for further uses in other systems, in an object
called `asmodee_outputs`; all objects are stored in */asmodee_outputs/rds/*:

* `summary`: summary of the ASMODEE results
* `results`: outputs of ASMODEE
* `plot_overall_deaths`: *ggplot2* object of the overall dynamics plot using
  death per capita on the y-axis
* `plot_overall_peaks`: *ggplot2* object of the overall dynamics plot using
  incidence as percentage of historical peak (excluding the last 90 days) on the
  y-axis


```{r }

asmodee_outputs <- list(
  summary = cases_outliers,
  results = cases_res,
  plot_overall_deaths = plot_overall_deaths,
  plot_overall_peaks = plot_overall_peaks
)

## export asmodee summary
file_name <- sprintf("asmodee_outputs_%s_%s.rds",
                     paste(params$who_region, collapse = "_"),
                     data_stamp)
path_out <- here::here("asmodee_outputs", "rds", file_name)
rio::export(asmodee_outputs, path_out)

## make a copy as latest file
latest_file_name <- sprintf("asmodee_outputs_%s_latest.rds",
                            paste(params$who_region, collapse = "_")
                            )
latest_path_out <- here::here("asmodee_outputs", "rds", latest_file_name)
file.copy(from = path_out, to = latest_path_out, overwrite = TRUE)

  
```



## Export main figures

Here we save the two key figures showing 'overall dynamics' to *png* files in
*asmodee_outputs/png/*:

```{r }

# Deaths PC as y axis
fig_name <- sprintf("waffle_deaths_%s_%s.png",
                    paste(params$who_region, collapse = "_"),
                    data_stamp)
fig_path <- here::here("asmodee_outputs", "png", fig_name)

png(fig_path, width = 22, height = 18, units = "cm", res = 150)
print(plot_overall_deaths)
dev.off()

# Make a copy as 'latest'
latest_fig_name <- sprintf("waffle_deaths_%s_latest.png",
                           paste(params$who_region, collapse = "_")
                           )
latest_fig_path <- here::here("asmodee_outputs", "png", latest_fig_name)
file.copy(from = fig_path, to = latest_fig_path, overwrite = TRUE)


# cases as % historical peak as y axis
fig_name <- sprintf("waffle_peaks_%s_%s.png",
                    paste(params$who_region, collapse = "_"),
                    data_stamp)
fig_path <- here::here("asmodee_outputs", "png", fig_name)

png(fig_path, width = 22, height = 18, units = "cm", res = 150)
print(plot_overall_peaks)
dev.off()

# Make a copy as 'latest'
latest_fig_name <- sprintf("waffle_peaks_%s_latest.png",
                           paste(params$who_region, collapse = "_")
                           )
latest_fig_path <- here::here("asmodee_outputs", "png", latest_fig_name)
file.copy(from = fig_path, to = latest_fig_path, overwrite = TRUE)

```



## Export notes

```{r }

if (is.null(notes)) {
  notes <- "Nothing to report. All countries were successfully included in the analyses."
}

## export asmodee summary
file_name <- sprintf("analysis_notes_%s_%s.md",
                     paste(params$who_region, collapse = "_"),
                     data_stamp)
path_out <- here::here("asmodee_outputs", "notes", file_name)
writeLines(notes, path_out, sep = "\n<br>\n")

## make a copy as latest file
latest_file_name <- sprintf("analysis_notes_%s_latest.md",
                            paste(params$who_region, collapse = "_")
                            )
latest_path_out <- here::here("asmodee_outputs", "notes", latest_file_name)
file.copy(from = path_out, to = latest_path_out, overwrite = TRUE)

```
